---
title: "DrugCentral MoA targets"
output:
  html_document:
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

# DrugCentral MoA targets with associated indications and diseases.

DrugCentral targets are curated from referenced publications and others sources. Mechanism-of-action (MoA) targets are convincingly
elucidated and published as such with referenced PubMed IDs. Together with approved indications, interpretable or associated with
diseases, these comprise very high confidence disease-to-target associations.

```{r echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "..")
base::date()
```

```{r message=FALSE, warning=FALSE}
library(RPostgreSQL, quietly = T)
library(readr, quietly=T)
library(data.table, quietly=T)
library(plotly, quietly=T)
```

## Connect to db; check db version.

```{r}
t0 <- proc.time()
dbcon <- dbConnect(PostgreSQL(), host="unmtid-dbs.net", port=5433, dbname="drugcentral", user="drugman", password="dosage")
```

```{r}
dbver <- dbGetQuery(dbcon, "SELECT * FROM dbversion", colClasses="character")
print(sprintf("DrugCentral version: %s (%s)\n", dbver$version[1], dbver$dtime[1]))
```

## Query db

This SQL joins tables with bioactivities, drug structures, and indications.

```{r}
sql <- read_file("sql/indication_targets.sql")
tgts <- dbGetQuery(dbcon, sql, colClasses="character")
setDT(tgts)
tgts[, moa := as.logical(!is.na(moa))]
```

### All indication-drug-target associations

```{r}
print(sprintf("Total indications: %d; drugs: %d; targets: %d; ind-tgt-pairs: %d", 
                tgts[, uniqueN(umls_cui)], tgts[, uniqueN(struct_id)], tgts[, uniqueN(gene)], nrow(unique(tgts[, .(umls_cui, gene)]))))
```

```{r}
act_sources <- tgts[, .(.N), by="act_source"]
plot_ly(type="pie", values=act_sources$N, labels=paste0(act_sources$act_source, "<br>(", act_sources$N, ")"), textinfo="label+percent", textfont=list(size=12)) %>%
  layout(title="DrugCentral target activity sources",
         margin = list(t=100, b=100),
         font=list(family="monospace", size=18), showlegend=T)
```

### MoA indication-drug-target associations

```{r}
print(sprintf("Total MOA indications: %d; drugs: %d; targets: %d; ind-tgt-pairs: %d", 
                tgts[(moa), uniqueN(umls_cui)], tgts[(moa), uniqueN(struct_id)], tgts[(moa), uniqueN(gene)], nrow(unique(tgts[(moa), .(umls_cui, gene)]))))
```

## Most-drugged indication-gene pairs

```{r}
tgts_moa <- unique(tgts[(moa) & !is.na(gene), .(n_drug = uniqueN(struct_id)), by = c("omop_concept", "umls_cui", "gene", "target_id", "target_name")])[order(omop_concept, gene)]
knitr::kable(tgts_moa[order(-n_drug)][1:25, .(omop_concept, gene, target_name, n_drug)], caption="Most-drugged indication-gene pairs")
```

## Most-drugged genes

```{r}
knitr::kable(tgts_moa[, .(n_drug = sum(n_drug)), by=c("gene", "target_name")][order(-n_drug)][1:25], caption="Most-drugged genes")
```


```{r}
write_delim(unique(tgts[, .(target_id)]), "data/dc_moa_targets.tgt_id")
write_delim(unique(tgts[, .(gene)]), "data/dc_moa_targets.gsymb")
write_delim(unique(tgts_moa[, .(umls_cui)]), "data/dc_moa_indications.cui")
```


```{r}
ok <- dbDisconnect(dbcon)
message(sprintf("elapsed time (total): %.2fs",(proc.time()-t0)[3]))
```
